---
title: "p8105_hw2_zw2974"
author: "Zihan Wu"
date: "9/27/2023"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
```

# Problem 1

## Importing and cleaning data tables

## For pols-month.csv

```{r}
polsmonth = read_csv("data/pols-month.csv") 
polsmonth_clean = polsmonth |> 
  janitor::clean_names()  |> 
  separate(mon, into = c("year", "month", "day"), sep = "-") |> 
  mutate(month = replace(month, month == "01", "January"))  |> 
  mutate(month = replace(month, month == "02", "February"))  |> 
  mutate(month = replace(month, month == "03", "March"))  |> 
  mutate(month = replace(month, month == "04", "April"))  |> 
  mutate(month = replace(month, month == "05", "May"))  |> 
  mutate(month = replace(month, month == "06", "June"))  |> 
  mutate(month = replace(month, month == "07", "July"))  |> 
  mutate(month = replace(month, month == "08", "August"))  |> 
  mutate(month = replace(month, month == "09", "September"))  |> 
  mutate(month = replace(month, month == "10", "October"))  |> 
  mutate(month = replace(month, month == "11", "November"))  |> 
  mutate(month = replace(month, month == "12", "December"))  |> 
  pivot_longer(cols = c("prez_dem", "prez_gop"), 
               names_to = "president",
               values_to = "value") |>
  filter(value == 1) |>
  mutate(president = case_when(
    president == "prez_gop" ~ "gop",
    president == "prez_dem" ~ "dem"
  )) |>
  select(-value, -day)
```


## For snp.csv

```{r}
snp = read_csv("data/snp.csv") 
snp_clean = snp |> 
  janitor::clean_names()  |> 
  separate(date, into = c("month", "day", "year"), sep = "/") |> 
  mutate(year = case_when(
                 nchar(year) == 1 ~ paste0("200", year),
                 nchar(year) == 2 & as.integer(year) <= 23 ~ paste0("20", year),
                 nchar(year) == 2 & as.integer(year) > 23 ~ paste0("19", year),
                 TRUE ~ year
             )) |> 
  mutate(month = factor(month, levels = 1:12, labels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))) |> 
  select(year, month, close)
```

## For unemployment.csv

```{r}
unemployment = read_csv("data/unemployment.csv")
unemployment_clean = unemployment |> 
  janitor::clean_names() |> 
  pivot_longer(cols = jan:dec, names_to = "month", values_to = "unemployment_rate") |> 
  mutate(month = replace(month, month == "jan", "January"))  |> 
  mutate(month = replace(month, month == "feb", "February"))  |> 
  mutate(month = replace(month, month == "mar", "March"))  |> 
  mutate(month = replace(month, month == "apr", "April"))  |> 
  mutate(month = replace(month, month == "may", "May"))  |> 
  mutate(month = replace(month, month == "jun", "June"))  |> 
  mutate(month = replace(month, month == "jul", "July"))  |> 
  mutate(month = replace(month, month == "aug", "August"))  |> 
  mutate(month = replace(month, month == "sep", "September"))  |> 
  mutate(month = replace(month, month == "oct", "October"))  |> 
  mutate(month = replace(month, month == "nov", "November"))  |> 
  mutate(month = replace(month, month == "dec", "December")) |> 
  select(year, month, unemployment_rate)
```

# Problem 2

## Read and clean the Mr. Trash Wheel sheet:

```{r}
mr_trash_wheel = read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = weight_tons * 500 /30) |> 
  mutate(type = "Mr.") |> 
  mutate(year = as.integer(year)) |> 
  drop_na(dumpster)
```

## Read and clean the Professor Trash Wheel sheet:

```{r}
professor_df = read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = weight_tons * 500 /30) |> 
  mutate(type = "Professor") |> 
  mutate(year = as.integer(year)) |> 
  drop_na(dumpster)

```

## Read and clean the Gwynnda Trash Wheel sheet:

```{r}
gwynnda_df = read_excel("data/202207 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:K108") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = weight_tons * 500 /30) |>
  mutate(type = "Gwynnda") |> 
  mutate(year = as.integer(year)) |>
  drop_na(dumpster)
```

## Read and clean the Professor Trash Wheel sheet:

```{r}
combination = bind_rows(mr_trash_wheel, professor_df, gwynnda_df) |> 
  relocate(type)

combination
```

## Write a paragraph about these data:

Number of observations: `r nrow(combination)`

Numbers of variables: `r ncol(combination)`

Names of key variables: `r names(combination)`

Total weight of trash collected by Professor Trash Wheel: `r sum(professor_df$weight_tons)`

Total number of cigarette butts collected by Gwynnda in July of 2021: `r sum(gwynnda_df$cigarette_butts[which(gwynnda_df$month == "July" & gwynnda_df$year == 2021)])`

# Problem 3

